---
title: "coursera_ML"
author: "monique-97"
date: "22 02 2021"
output:
  pdf_document: default
  html_document: default
---

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

**data source:**
Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

# aim of the project
The goal of your project is to predict the manner of exercises, this is the "classe" variable in the training set. 




# read libraries
```{r, echo = TRUE}
library(caret)
library(randomForest)
library(kernlab)
library(rpart.plot)

```


# read data
```{r, echo = TRUE}
train_data <- read.csv("pml-training.csv")
test_data <- read.csv("pml-testing.csv")
```
We have 19622 observations and 160 variables in train data, and 20 observations and 160 variables in test data.

# data pre-processing

## remove NAs
```{r, echo = TRUE}
train_data <- train_data[, colSums(is.na(train_data)) == 0] 
test_data <- test_data[, colSums(is.na(test_data)) == 0] 
``` 


## take only columns of interest
```{r, echo = TRUE}
training <- train_data[,-c(1:7)]
testing <-test_data[,-c(1:7)]
```

## change values into numeric
```{r, echo = TRUE}
classe <- train_data$classe
train_clean <- train_data[, sapply(train_data, is.numeric)]
train_clean$classe <- classe
test_clean <- test_data[, sapply(test_data, is.numeric)]
```

## split training data into a training dataset and a data for validation, set seed to get reproducible results
```{r, echo = TRUE}
set.seed(199997)
training_part <- createDataPartition(train_clean$classe, p=0.65, list=FALSE)
train_values <- train_clean[training_part, ]
test_values <- train_clean[-training_part, ]
```

## data visualization: frequency of each level
```{r, echo = TRUE}
barplot(table(train_values$classe), col="blue", main="Frequency of each level", xlab=" ", ylab="frequency")
```
Level A is the most frequent.

# builting the model
Random forest prediction model will be applied.

## run random forest
```{r, echo = TRUE}
controltr <- trainControl(method="cv", 5)
model_random <- train(classe ~ ., data=train_values, method="rf", trControl=controltr, ntree=100)
model_random
```

# cross validation
## estimate model performance
```{r, echo = TRUE}
prediction <- predict(model_random, test_values)
confusionMatrix(test_values$classe, prediction)
```

# out of sample error
## calculate the expected out of sample error and accuracy
```{r, echo = TRUE}
error <- 1 - as.numeric(confusionMatrix(test_values$classe, prediction)$overall[1])
acc <- postResample(prediction, test_values$classe)
```

Out-of-sample error is 0.015%.

## apply model to test dataset
```{r, echo = TRUE}
final_result <- predict(model_random, test_clean[, -length(names(test_clean))])
```


Machine learning algorithm was applied to the 20 test cases available in the test data.

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. 


